Open Questions for Computer Science and Cheminformatics

### 2011-11-20 OpenQuestions.tex
Minor rewrite
Another challenge was added

One key concept of the linked data web, the semantic web, is that different data sources can be readily integrated with each other. Still, in the field
of Cheminformatics, we are not only interest in linking two molecules 
(the linking normalization problem for different protomers, tautomers, or special cases of isomerisms remain open), but we
are also interested in being able to search efficiently within molecules when being linked via semantic web approaches.
What could be algorithmic solutions for this?
\end{enumerate}

@@

\section*{Data mining}
\begin{enumerate}
\item evaluation of similarities in a heterogeneous network (JKW: What are heterogenous networks?).  What is a specific example here?
\item integrate chemical structure information with ontologies.  Specific example problem?
\item Classify molecular descriptors up to equivalence, and dependence on one another.
\item systems-level understanding of small molecules (JKW: which systems? This will help being clearer on the challenge).  What does this mean, and what would a specific challenge problem be?
%\item \emph{Efficient molecule browsing, e.g. on scaffold level}.
%
%Chemical Abstract Services have a molecule browsing tool called SubScape, which allows to browse large-scale
%chemical spaces efficiently. What could be large-scale solutions for doing this within (combined and aligned) 
%public databases.
%
\item \emph{Large-scale browsing of molecular property spaces, e.g. on scaffold level, side-effect-level, ...}.

Certain molecules might have hundreds of biological activities, side-effects in humans (SIDER database \cite{Kuhn_Campillos_Letunic_Jensen_Bork_2010}), or
many other properties attached to them. What are large-scale mining and visulization options? 
How can we mine private and public data sources at the very same time?
%
\item \emph{Patent text mining (curation)}.

@@

Let us assume an organization, e.g. a commercial company, has a single or a small number of established KM and ELN products. 
How can we improve the maintenance, leveraging, and collaboration with many external partners (each of them potentially
with another KM/ELN solution)? Which party is hosting which data in which data structure (ontologies?), and 
how can we ensure that only pre-defined data entries 
(and a limited number of annotations, e.g. bioological activities) are visible to a partner.
How can this be organized for a multitude of partners? Cloud computing, user management, encryption granularity and efficient security management?
\end{enumerate}

>>>>

One key concept of the linked data web, the semantic web, is that different data sources can be readily integrated with each other. Still, in the field
of Cheminformatics, we are not only interest in linking two molecules 
(the linking normalization problem for different protomers, tautomers, or special cases of isomerisms remain open), but we
are also interested in being able to search efficiently within molecules when being linked via semantic web approaches. Typical
searches will require being able to apply substructure or similarity searches.
What could be algorithmic solutions for this?
\end{enumerate}

@@

\section*{Data mining}
\begin{enumerate}
\item evaluation of similarities in a heterogeneous network (JKW: What are heterogenous networks?).  What is a specific example here?
\item integrate chemical structure information with ontologies.  Specific example problem?
\item Classify molecular descriptors up to equivalence, and dependence on one another.
\item systems-level understanding of small molecules (JKW: which systems? This will help being clearer on the challenge).  What does this mean, and what would a specific challenge problem be?
%\item \emph{Efficient molecule browsing, e.g. on scaffold level}.
%
%Chemical Abstract Services have a molecule browsing tool called SubScape, which allows to browse large-scale
%chemical spaces efficiently. What could be large-scale solutions for doing this within (combined and aligned) 
%public databases.
%
\item \emph{Dynamic similarity search on instant binary vectors}.

Binary feature vectors are a common practice for chemical similarity searches. The typical process starts with 1. creating
binary substructure (or other feature vectors) \cite{citeulike:8530538}, 2. creating fixed-length binary vectors of typically
1024 bits for reducing space requirements and speeding up further similarity searches (by loosing some accuracy), 3. creating 
further pre-computations for speeding up threshold based similarity searches \cite{doi:10.1021/ci800076s}.\\
Still, if we are interested to employ dynamic changes in the similarity encodings, e.g. using only a set of binary features,
then previously done hashing or pre-computations might need to be redone efficiently on the fly. Finally, the major goal
is to employ similarity searches \cite{doi:10.1021/ci200235e} on a scale of multiple million entries and more optimizations and
benchmarking studies are urgently required, e.g. using GPUs
\cite{doi:10.1021/ci1004948}, optimizing pair-wise similarity calculations \cite{MINF:MINF201100050}.
%
\item \emph{Large-scale browsing of molecular property spaces, e.g. on scaffold level, side-effect-level, ...}.

Certain molecules might have hundreds of biological activities, side-effects in humans (SIDER database \cite{Kuhn_Campillos_Letunic_Jensen_Bork_2010}), or
many other properties attached to them. What are large-scale mining and visulization options? 
How can we mine private and public data sources at the very same time?
%
\item \emph{Chemical image/text mining in patents (curation)}.

@@

Let us assume an organization, e.g. a commercial company, has a single or a small number of established KM and ELN products. 
How can we improve the maintenance, leveraging, and collaboration with many external partners (each of them potentially
with another KM/ELN solution)? Which party is hosting which data in which data structure (ontologies?), and 
how can we ensure that only pre-defined data entries 
(and a limited number of annotations, e.g. biological activities) are visible to a partner.
How can this be organized for a multitude of partners? Cloud computing, user management, encryption granularity and efficient security management?
\end{enumerate}
